---
title: "2019 OmegaQuant Plot v1.3"
author: "Benny Rice"
date: "03/26/2019"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

***
#### Notes

- Data types to visualize and/or analyze in this file:

1. OmegaQuant fatty acid results (using a sample of data from the CRS2017 project and the Kiribati 2020 project)

***
#### Defining thresholds

- See **[GitHub](https://github.com/bennyrice/OmegaQuant)** for associated documentation on cut-offs (i.e., Harris et al 2017)

##### Omega-3 index

```{r t1, eval = FALSE}
# Omega-3 index values >  6.1%            = no increased cardiovascular health risk
# Omega-3 index values <= 6.1% and > 4.0% = increased cardiovascular health risk
# Omega-3 index values <= 4.0%            = greatly increased cardiovascular health risk
```

##### Trans fat index

**Note:** Trans fat index seems to be missing from the Kiribati data

```{r t2, eval = FALSE}
# Trans fat index values <  1.00%             = no increased cardiovascular health risk
# Trans fat index values >= 1.00% and < 1.75% = intermediate cardiovascular health risk
# Trans fat index values >= 1.75%             = increased cardiovascular health risk
```
***

Set up: Defining main working directory
```{r i1, echo = FALSE, message = FALSE}
mainDir <- '/Users/blrice/Dropbox/Lab Projects/CRS - Data Cleaning/WORK/C 2019 Preliminary Analysis Work/20190128 OmegaQuant/20200729 Adjusting code for Kiribati Data/'
```

Set up: Loading packages (*not all may be necessary)
```{r i2, echo = FALSE, message = FALSE}
library(tidyverse)
library(knitr)
library(kableExtra)
library(ggthemes)
library(reshape2)
library(ggpubr)
library(waffle)
```

Set up: Reading in CSVs

- Excel sheets `MAHERY Results 01252019.xlsx` and `OmegaQuant Results - Harvard - Golden - P378-US1 - DBS - 16Jul2020.xlsx` manually saved as CSV files
- Subsequent analysis may need to link to age and sex data (not done here)

```{r i3, echo = FALSE, message = FALSE}
filename_df.CRS <- "MAHERY Results 01252019.csv"
#Skip 11 lines at the top of the CSV
df.CRS <- read.csv(file.path(mainDir, filename_df.CRS), skip = 11, stringsAsFactors = FALSE)
#Delete junk rows
df.CRS <- df.CRS %>% filter(!is.na(calculated_age)) %>% 
  #Keep ID, region, village, Omega 3 and trans fat columns
  select(OmQ_filter_paper_id, region_name, village_name, Omega.3.Index, Trans.Fat.Index) %>%
  #Remove % sign from columns, convert to numeric
  mutate(Omega.3.Index   = as.numeric(gsub("%", "", Omega.3.Index))) %>%
  mutate(Trans.Fat.Index = as.numeric(gsub("%", "", Trans.Fat.Index)))

filename_df.KIRI <- "OmegaQuant Results - Harvard - Golden - P378-US1 - DBS - 16Jul2020.csv"
#Skip 15 lines at the top of the CSV
df.KIRI <- read.csv(file.path(mainDir, filename_df.KIRI), skip = 15, stringsAsFactors = FALSE)
#Delete junk rows
df.KIRI <- df.KIRI %>% filter(Omega.3.Index != "") %>% 
  #Keep ID, Date, Location, Omega 3 columns
  select(Info.on.Collection.Card, Date, Location, Omega.3.Index) %>%
  #Remove % sign from columns, convert to numeric
  mutate(Omega.3.Index   = as.numeric(gsub("%", "", Omega.3.Index)))
```

***


***

<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">

# (1) Fatty Acid Data

</div>

***

***

#### (1.1) Applying Omega 3 thresholds:

***

(1.1.1) Specifying Omega 3 outcomes
```{r 1.1.1, eval = FALSE}
# Omega-3 index values >  6.1%            = no increased cardiovascular health risk
# Omega-3 index values <= 6.1% and > 4.0% = increased cardiovascular health risk
# Omega-3 index values <= 4.0%            = greatly increased cardiovascular health risk

# (OM3_1) Omega.3.Index >  6.1                       ~ '1_no_increased_risk'
# (OM3_2) Omega.3.Index <= 6.1 & Omega.3.Index > 4.0 ~ '2_increased_risk'
# (OM3_3) Omega.3.Index <= 4.0                       ~ '3_greatly_increased_risk'
```

(1.1.2) Filtering Omega 3 data per thresholds
```{r 1.1.2A, echo = FALSE}
df.CRS <- df.CRS %>% mutate(OM3_outcome_cat = case_when(
  Omega.3.Index >  6.1                       ~ '1_no_increased_risk',
  Omega.3.Index <= 6.1 & Omega.3.Index > 4.0 ~ '2_increased_risk',
  Omega.3.Index <= 4.0                       ~ '3_greatly_increased_risk'))
df.KIRI <- df.KIRI %>% mutate(OM3_outcome_cat = case_when(
  Omega.3.Index >  6.1                       ~ '1_no_increased_risk',
  Omega.3.Index <= 6.1 & Omega.3.Index > 4.0 ~ '2_increased_risk',
  Omega.3.Index <= 4.0                       ~ '3_greatly_increased_risk'))
```

Preview of cleaned and filtered Omega 3 data:
```{r 1.1.2B, echo = FALSE}
kable(df.CRS[1:100, ],  caption = "CRS: Preview of cleaned and filtered Omega 3 data")  %>% 
  kable_styling(bootstrap_options = c("striped", "hover"), font_size = 11) %>% scroll_box(width = "100%", height = "300px")

kable(df.KIRI[1:100, ], caption = "KIRI: Preview of cleaned and filtered Omega 3 data") %>% 
  kable_styling(bootstrap_options = c("striped", "hover"), font_size = 11) %>% scroll_box(width = "100%", height = "300px")
```

- (even here you can already see a difference between Madagascar and Kiribati)

***
***

#### (2.1) Applying Trans fat thresholds:

- (**NOTE:** No trans data available for Kiribati in the date file)

***

(2.1.1) Specifying Trans fat outcomes
```{r 2.1.1, eval = FALSE}
# Trans fat index values <  1.00%             = no increased cardiovascular health risk
# Trans fat index values >= 1.00% and < 1.75% = intermediate cardiovascular health risk
# Trans fat index values >= 1.75%             = increased cardiovascular health risk

# (TFAT_1) Trans.Fat.Index <  1.00                          ~ '1_no_increased_risk'
# (TFAT_2) Trans.Fat.Index >= 1.00 & Trans.Fat.Index < 1.75 ~ '2_increased_risk'
# (TFAT_3) Trans.Fat.Index >= 1.75                          ~ '3_greatly_increased_risk'
```

(2.1.2) Filtering Trans fat data per thresholds
```{r 2.1.2A, echo = FALSE}
df.CRS <- df.CRS %>% mutate(TFAT_outcome_cat = case_when(
  Trans.Fat.Index <  1.00                          ~ '1_no_increased_risk',
  Trans.Fat.Index >= 1.00 & Trans.Fat.Index < 1.75 ~ '2_increased_risk',
  Trans.Fat.Index >= 1.75                          ~ '3_greatly_increased_risk'))
```

Preview of cleaned and filtered Trans fat data:
```{r 2.1.2B, echo = FALSE}
kable(df.CRS[1:100, ],  caption = "CRS: Preview of cleaned and filtered Trans fat data")  %>% 
  kable_styling(bootstrap_options = c("striped", "hover"), font_size = 11) %>% scroll_box(width = "100%", height = "300px")
```

***

#### (2.3) Visualizing Omega 3 and Trans fat data

***

(2.3.1) MADAGASCAR (CRS): Proportions of the sample population with a given Omega 3 level
```{r 2.3.1, echo = FALSE, fig.height = 7}
df.CRS %>% group_by(OM3_outcome_cat) %>% summarize(n = length(Omega.3.Index), .groups = 'drop') %>% ungroup() %>%
  ggplot(aes(fill = OM3_outcome_cat, values = n)) +
  geom_waffle(color = "white", size = .25, n_rows = 20, flip = TRUE) +
  scale_x_discrete() + 
  scale_y_continuous(labels = function(x) x * 20, # make this multiplyer the same as n_rows
                     expand = c(0,0)) +
  ggthemes::scale_fill_tableau(name=NULL) +
  coord_equal() +
  labs(
    title = "Omega 3 Status",
    subtitle = "(CRS2017 Madagascar)",
    x = "Risk category",
    y = "Count"
  ) +
  theme_minimal() +
  theme(panel.grid = element_blank(), axis.ticks.y = element_line()) +
  guides(fill = guide_legend(reverse = TRUE))
```

***

***

(2.3.2) KIRIBATI: Proportions of the sample population with a given Omega 3 level
```{r 2.3.2, echo = FALSE, fig.height = 7}
df.KIRI %>% group_by(OM3_outcome_cat) %>% summarize(n = length(Omega.3.Index), .groups = 'drop') %>% ungroup() %>%
  ggplot(aes(fill = OM3_outcome_cat, values = n)) +
  geom_waffle(color = "white", size = .25, n_rows = 20, flip = TRUE) +
  scale_x_discrete() + 
  scale_y_continuous(labels = function(x) x * 20, # make this multiplyer the same as n_rows
                     expand = c(0,0)) +
  ggthemes::scale_fill_tableau(name=NULL) +
  coord_equal() +
  labs(
    title = "Omega 3 Status",
    subtitle = "(Kiribati)",
    x = "Risk category",
    y = "Count"
  ) +
  theme_minimal() +
  theme(panel.grid = element_blank(), axis.ticks.y = element_line()) +
  guides(fill = guide_legend(reverse = TRUE))
```

***

***

(2.3.3) MADAGASCAR (CRS): Proportions of the sample population with a given Trans fat level
```{r 2.3.3, echo = FALSE, fig.height = 7}
df.CRS %>% group_by(TFAT_outcome_cat) %>% summarize(n = length(Trans.Fat.Index), .groups = 'drop') %>% ungroup() %>%
  ggplot(aes(fill = TFAT_outcome_cat, values = n)) +
  geom_waffle(color = "white", size = .25, n_rows = 20, flip = TRUE) +
  scale_x_discrete() + 
  scale_y_continuous(labels = function(x) x * 20, # make this multiplyer the same as n_rows
                     expand = c(0,0)) +
  ggthemes::scale_fill_tableau(name=NULL) +
  coord_equal() +
  labs(
    title = "Trans Fat Status",
    subtitle = "(CRS2017 Madagascar)",
    x = "Risk category",
    y = "Count"
  ) +
  theme_minimal() +
  theme(panel.grid = element_blank(), axis.ticks.y = element_line()) +
  guides(fill = guide_legend(reverse = TRUE))
```

***

***

(2.3.4) **MADAGASCAR (CRS2017)**: Scatter plot of Omega 3 Levels
```{r 2.3.4, echo = FALSE, fig.height = 7}
df.CRS %>% group_by(village_name) %>% arrange(OmQ_filter_paper_id) %>% 
  mutate(row_index = 1:length(Omega.3.Index)) %>% ungroup() %>%
  mutate(OM3_outcome_cat = 
           factor(OM3_outcome_cat, 
                  levels = c("1_no_increased_risk", "2_increased_risk", "3_greatly_increased_risk"))) %>%
  ggplot(aes(x = row_index, y = Omega.3.Index, color = OM3_outcome_cat)) +
  geom_point() +
  scale_color_manual(values=c("#4e6cef", "#ffa000", "#dd2c00")) +
  facet_grid(cols = vars(village_name), scales = "free_x") +
  theme_clean() + 
  ylim(0, 17) +
  theme(legend.position = "top", legend.text = element_text(size = 5),
        axis.text.x = element_blank(), axis.title.x = element_blank(), axis.ticks.x = element_blank(),
        strip.text = element_text(size = 5))
```

(2.3.5) **KIRIBATI**: Scatter plot of Omega 3 Levels
```{r 2.3.5, echo = FALSE, fig.height = 7}
df.KIRI %>% mutate(row_index = 1:length(Omega.3.Index)) %>%
  mutate(OM3_outcome_cat = 
           factor(OM3_outcome_cat, 
                  levels = c("1_no_increased_risk", "2_increased_risk", "3_greatly_increased_risk"))) %>%
  ggplot(aes(x = row_index, y = Omega.3.Index, color = OM3_outcome_cat)) +
  geom_point() +
  scale_color_manual(values=c("#4e6cef", "#ffa000", "#dd2c00")) +
  theme_clean() + 
  ylim(0, 17) +
  theme(legend.position = "top", legend.text = element_text(size = 5),
        axis.text.x = element_blank(), axis.title.x = element_blank(), axis.ticks.x = element_blank())

```

***


